Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-09-08 paper/main.tex
sections for measurment transforms

We propose that the map $\pi$ can express the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, as discussed in \autoref{sec:atct:fb:base}. For our purposes, the advantage of fiber bundles is that the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every index point has a corresponding field type and the field types must be the same for every index point. We propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner, as discussed in \autoref{sec:atct:fb:sections}.

@@

\subsubsection{Examples}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{trivnon.png}
  \caption{A fiber bundle with a cylinderical total space, which is a trivial bundle, and bundle with a mobius band total space, which is non-trivial, can be covered by the same set of local trivializations. For example, they can both be decomposed into two planes over two intervals $\left(-\epsilon,\pi+\epsilon\right), \left(\pi-\epsilon, 2\pi+\epsilon\right)$. The transition maps that transforms this set of local trivializations into a cylinder are two identity map aligning two \textit{up} fibers, while the transition maps that transform this set into a mobius bundle are an identity map and a map that aligns an \textit{up} fiber with a \textit{down} fiber.  \note{in final may change this to something that's not half half to make the arbitraryness more clear/break it up into smaller pieces} \label{fig:atct:nontrivialbundle}}
\end{figure}

As shown in \autoref{fig:atct:nontrivialbundle}, a trivial bundle with a cylinderical total space and a non-trivial bundle with mobius band total spacee can be covered with the same local trivializations; the distinguishing factor between these bundles is that the transition maps on these trivializations differ. The fibers in the cylinder bundle align in the same direction, \textit{up}, such that the transition maps that assemble the local trivializations into the cylinderical total space are all identity maps alinging two \textit{up} fibers. In the mobius bundle, there is twist where the fibers align in different directions, shown in the inset in \autoref{fig:atct:nontrivialbundle}. This twist is not present in either local trivialization because, by definition, fibers need to be aligned in local trivializations; therefore the transition map that aligns the fiber at the overlap between $\left(-\epsilon, \epsilon\right)$ must account for this flip by aligning \textit{up} fibers \textit{down} fibers. In this example, the transition map on the $\left(\pi-\epsilon, \pi+\epsilon\right)$ is an identity map since the fibers are aligned in the same direction on that section of the mobius band. \note{add something in the figure highlighting this part of the cylinder and the mobius band}

@@

\paragraph{Transform Continuity}\\

The base space transformation is a pointwise continuous map from one open set to another open set in the same base space
\begin{equation}
  \label{eq:atct:morphism:base}
 \dfunch: \openset^{\prime}\rightarrow \openset
 \end{equation}
such that $\openset, \openset^{\prime} \subseteq \dbase$. This means $\openset$ and $\openset^{\prime}$ are of the same topology type. To correctly align the sections with the remapped base space, there is a a corresponding section pullback function
\begin{equation}
  \label{eq:atct:morphism:basepull}
  \dfuncpull \dsection \restriction_{\openset^{\prime}}: \dsection\restriction_{\openset^{\prime}} \mapsto \dsection \restriction_{\openset^{\prime} \circ \dfunch} 
\end{equation}
such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{\prime})}$. This means that the base space transformation \dfunch\ does not change the data values at a given point
\begin{equation}
  \label{eq:atct:morphism:verify_base}
  \dsection(\dbase) = \dfuncpull\dsection(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbasepoint^{\prime}))
\end{equation} 
where $\dfunch(\dbasepoint^{\prime}) = \dfunch(\dbasepoint)$.

\paragraph{Transform Records}\\

As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a monoid action on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \label{eq:atct:morphism:fiber}
  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}. Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.

\paragraph{Transform Continuity and Records}

We define a full data transformation as one that induces both a remapping of the index space and a change in the data values
\begin{equation}
  \label{eq:atct:morphism:all}
  \dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc} \circ \dfunchc
\end{equation}
which gives us an equation that can express transformations that have both a base space change and a fiber change.

@@

\section{Discussion: Feasibility of Math as Design Spec}
\label{sec:discussion}
\begin{figure}
  \includegraphics[width=1\columnwidth]{full_line.png}
  \caption{\label{fig:construction:artist:combo}}
\end{figure}
The framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components. In this section, we introduce a toy example of building an artist out of the components introduced in \autoref{sec:construction} to illustrate how components that adhere to these specifications are maintainable, extendible, scalable, and support concurrency.

>>>>

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{trivnon.png}
  \caption{A fiber bundle with a cylinderical total space, which is a trivial bundle, and bundle with a mobius band total space, which is non-trivial, can be covered by the same set of local trivializations. For example, they can both be decomposed into two planes over two intervals $\left(-\epsilon,\pi+\epsilon\right), \left(\pi-\epsilon, 2\pi+\epsilon\right)$. The transition maps that transforms this set of local trivializations into a cylinder are two identity map aligning two \textit{up} fibers, while the transition maps that transform this set into a mobius bundle are an identity map and a map that aligns an \textit{up} fiber with a \textit{down} fiber.  \note{in final may change this to something that's not half half to make the arbitraryness more clear/break it up into smaller pieces} \label{fig:atct:nontrivialbundle}}
\end{figure}

As shown in \autoref{fig:atct:nontrivialbundle}, a trivial bundle with a cylinderical total space and a non-trivial bundle with mobius band total spacee can be covered with the same local trivializations; the distinguishing factor between these bundles is that the transition maps on these trivializations differ. The fibers in the cylinder bundle align in the same direction, \textit{up}, such that the transition maps that assemble the local trivializations into the cylinderical total space are all identity maps alinging two \textit{up} fibers. In the mobius bundle, there is twist where the fibers align in different directions, shown in the inset in \autoref{fig:atct:nontrivialbundle}. This twist is not present in either local trivialization because, by definition, fibers need to be aligned in local trivializations; therefore the transition map that aligns the fiber at the overlap between $\left(-\epsilon, \epsilon\right)$ must account for this flip by aligning \textit{up} fibers \textit{down} fibers. In this example, the transition map on the $\left(\pi-\epsilon, \pi+\epsilon\right)$ is an identity map since the fibers are aligned in the same direction on that section of the mobius band. \note{add something in the figure highlighting this part of the cylinder and the mobius band, possibly add in a section}

We propose that the total space of a bundle can encode the mathematical space in which a  dataset is embedded, the base space can encode the topological properties of the dataset, and the fiber space can encode the data types of the record fields of the dataset. The map $\pi$ expresses the formal binding between having a typed data field, discussed in \autoref{sec:atct:fb:fiber}, and a corresponding point in the topological structure, which is discussed in \autoref{sec:atct:fb:base}. Fiber bundles are a good abstract data representation for visualization because the field type and topological structure are unconstrained in terms of dimensionality and the only conditions that must be satisfied are that every point in the base space has a corresponding field of values and the field types must be the same for every point in the base space. We propose that modeling data as sections provides a way to encapsulate topological and field structure in a uniform dimension and type independent manner, as discussed in \autoref{sec:atct:fb:sections}.

@@

\subsubsection{Example}

@@

\paragraph{\textbf{Topological structure}}
\noindent
The base space transformation is a pointwise continuous map from one open set to another open set in the same base space
\begin{equation}
  \label{eq:atct:morphism:base}
 \dfunch: \openset^{\prime}\rightarrow \openset
 \end{equation}
such that $\openset, \openset^{\prime} \subseteq \dbase$. This means $\openset$ and $\openset^{\prime}$ are of the same topology type. To correctly align the sections with the remapped base space, there is a a corresponding section pullback function
\begin{equation}
  \label{eq:atct:morphism:basepull}
  \dfuncpull \dsection \restriction_{\openset^{\prime}}: \dsection\restriction_{\openset^{\prime}} \mapsto \dsection \restriction_{\openset^{\prime} \circ \dfunch} 
\end{equation}
such that $\dsection|_{\openset} = \dfuncpull\dsection|_{\opensetc^{\prime}}$ because $\dsection|_{\openset} = \dsection|_{\dfunch(\opensetc^{\prime})}$. This means that the base space transformation \dfunch\ does not change the data values at a given point
\begin{equation}
  \label{eq:atct:morphism:verify_base}
  \dsection(\dbase) = \dfuncpull\dsection(\dbasepoint^{\prime}) = \dsection(\dfunch(\dbasepoint^{\prime}))
\end{equation} 
where $\dfunch(\dbasepoint^{\prime}) = \dfunch(\dbasepoint)$.

\paragraph{\textbf{Records}}
\noindent
As introduced in \autoref{eq:artist:sheaves:monoid_morphism}, the fiber transformation $\dfunct: \dfuncpull \dtotal_{\dbasepoint^{\prime}} \rightarrow \dfuncpull \dtotal_{\dbasepoint^{\prime}}$ is a monoid action on the fiber $\dfunct \in Hom(\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}},\dfuncpull\dfiber\restriction_{\dbasepoint^{\prime}})$ restricted to a point $\dbasepoint^{\prime} \in \openset^{\prime}$. The fiber transformation is a change in section 
\begin{equation}
  \label{eq:atct:morphism:fiber}
  \dfunct: \dfuncpull \dsection \restriction_{\openset^{\prime}} \mapsto \dfuncpull \dsection^{\prime} \restriction_{\openset}
\end{equation}
where $\dsection, \dsection^{\prime} \in \Gamma(\openset^{\prime}, \dfuncpull\dtotal\restriction_{\openset^{\prime}})$. Since there are many different fiber transformations,  there is not a uniform way of expressing how this morphism acts on sections unlike in \autoref{eq:atct:morphism:verify_base}. Examples of a fiber transformation are actions such as partial ordering for ranking purpose \cite{bruggemannRankingPrioritizationMultiindicator2011} and the permutation, ordering, translation, and scaling codified as Steven's measurement scales \cite{stevensTheoryScalesMeasurement1946, leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}.

\paragraph{\textbf{Topological structure and records}} 
\noindent We define a full data transformation as one that induces both a remapping of the index space and a change in the data values
\begin{equation}
  \label{eq:atct:morphism:all}
  \dfuncc: \dsectionc\restriction_{\opensetc} \mapsto \dsectionc^{\prime}\restriction_{\opensetc} \circ \dfunchc
\end{equation}
which gives us an equation that can express transformations that have both a base space change and a fiber change.

@@

\section{Discussion: Feasibility as Design Spec}
\label{sec:discussion}
\begin{figure}
  \includegraphics[width=1\columnwidth]{full_line.png}
  \caption{\label{fig:construction:artist:combo}}
\end{figure}
The framework specified in \autoref{sec:artist} and \autoref{sec:construction} describes how to build structure preserving visualization components, but it is left to the library developer to follow these guidelines when building and reusing components. In this section, we introduce a toy example of building an artist out of the components introduced in \autoref{sec:construction} to illustrate how components that adhere to these specifications are maintainable, extendible, scalable, and support concurrency.
